The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley Long Wavelength Array

### 2017-10-20 paper.tex
add coauthors and begin to address comments

@@

\newcommand{\swinburne}{
    \affiliation{
        Centre for Astrophysics \& Supercomputing,
        Swinburne University of Technology,
        PO Box 218,
        Hawthorn, VIC 3122, Australia;
    }
}

@@

@@

% alphabetical by last name
\author{Steven~W.~Ellingson}  \vt
\author{Lincoln~J.~Greenhill} \harvard
\author{T.~Joseph~W.~Lazio}   \jpl
\author{Daniel~C.~Price}      \harvard \berkeley \swinburne
\author{David~P.~Woody}       \ovro

%\author{Hugh Garsden}
%\affiliation{Harvard}
%
%\author{Stephen Bourke}
%\affiliation{Chalmers}
%
%\author{Esayas B. Shume} % contributed ionosphere text
%\affiliation{Jet Propulsion Laboratory, 4800 Oak Grove Dr, Pasadena, CA 91109}

@@

\begin{abstract}
    A host of new low-frequency radio telescopes are attempting to detect the 21~cm transition at
    cosmological distances. These measurements hold promise to directly probe star and galaxy
    formation at redshifts $20 \gtrsim z \gtrsim 7$, but are limited by the dynamic range they can
    achieve against foreground sources of low-frequency radio emission. Consequently, there is a
    growing demand for modern, high-fidelity maps of the sky at frequencies below 200 MHz for use in
    foreground modeling and removal.  We use the Owens Valley Long Wavelength Array (OVRO-LWA) to
    generate 8 new maps of the sky north of $\delta=-30^\circ$ with 15 arcmin angular resolution and
    at frequencies evenly spaced between 36.528~MHz and 73.152~MHz with 24~kHz bandwidth.  Existing
    full-sky maps at these frequencies have angular resolutions $\ge 2^\circ$ and so these maps are
    a 10-fold improvement in angular resolution.  The thermal noise in each map is $\sim$800
    mJy/beam.  Future iterations will incorporate total power radiometry, and additional bandwidth
    and integration time. This will improve the thermal noise while simultaneously pushing to higher
    angular resolutions with the planned expansion of the OVRO-LWA to 2.6~km baselines. These maps
    are generated from the application of a new widefield imaging technique for drift-scanning
    interferometers, Tikhonov-regularized $m$-mode analysis imaging. This algorithm constructs
    images of the entire sky in a single synthesis imaging step with exact treatment of widefield
    effects. We additionally introduce this technique and describe how CLEAN can be adapted to
    deconvolve maps generated by $m$-mode analysis imaging.  These maps serve as a first step on the
    path to the use of more sophisticated foreground filters in 21~cm cosmology incorporating the
    measured angular and frequency structure of all foreground contaminants.
\end{abstract}

@@

At redshifts $20 \gtrsim z \gtrsim 7$ the 21~cm hyperfine structure line of neutral hydrogen is
expected to produce a 10 to 100 mK perturbation in the Cosmic Microwave Background (CMB) spectrum
\citep{2006PhR...433..181F, 2012RPPh...75h6901P}. The amplitude of this perturbation on a given
line-of-sight is a function of the neutral fraction of hydrogen, the baryon overdensity, the spin
temperature relative to the CMB temperature at the given redshift, and the line-of-sight peculiar
velocity of the gas.  The spatial power spectrum of this perturbation is thought to be dominated by
inhomogeneous heating of the IGM at $z\sim 20$ \citep{2014MNRAS.437L..36F}, and by growing ionized
bubbles during the EoR at $z\sim 7$ where a detection can constrain the ionizing efficiency of early
galaxies, the UV photon mean-free-path, and the minimum halo mass that can support star formation
\citep{2015MNRAS.449.4246G}.

Current 21~cm cosmology experiments can be broadly separated into two classes: global signal
experiments that are aiming to detect the spectral signature of the cosmologically redshifted 21~cm
transition after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21~cm perturbations.  Ongoing global signal experiments include EDGES
\citep{2010Natur.468..796B, 2017ApJ...847...64M}, LEDA \citep{2017arXiv170909313P}, BIGHORNS
\citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2
\citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21~cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2008AJ....136..641R} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.5}\,.
\end{equation}
Therefore experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. However the low-frequency radio sky is composed of several
components: galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission
and absorption from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally a
foreground removal strategy should be informed by the measured spatial structure and frequency
spectrum of all foreground components.  This possibility is limited by the availability of suitable
high-fidelity low-frequency sky maps on angular scales ranging from tens of degrees to arcminutes.

The Global Sky Model (GSM) \citep{2008MNRAS.388..247D, 2017MNRAS.464.3486Z} is currently the most
commonly used foreground model. The GSM is a data-driven interpolation of various maps between 10
MHz and 100 GHz. However the majority of information contained in the GSM is derived at frequencies
$>1.4$ GHz where the majority of the modern, high-fidelity input maps are located.  Below 408 MHz,
the GSM is a strong function of the Haslam 408 MHz map \citep{1981A&A...100..209H,
1982A&AS...47....1H, 2015MNRAS.451.4311R}. As a consequence, the GSM is not a faithful
representation of the sky below 200 MHz. For instance, \citet{2017MNRAS.469.4537D} report errors of
$\pm 50\%$ between the GSM and their own maps at 74 MHz. This is largely due to the increasing
contribution of free-free absorption and modifications to the synchrotron spectral index at low
frequencies.

@@

This paper presents a series of new low-frequency maps of the sky between 36.528~MHz and 73.152~MHz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21~cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. We anticipate additional anciliary science taking advantage of the combination
of high fidelity and high resolution of these maps, including but not limited to studies of the
cosmic ray emissivity at low frequencies, searches for giant radio galaxies, and constraining the
galactic synchrotron spectrum. The maps will be made freely available online at the Legacy Archive
for Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

@@

The goal of all imaging algorithms is to estimate the brightness of the sky $I_\nu(\hat r)$ in the
direction $\hat r$ and frequency $\nu$.  A radio interferometer measures the visibilities
$V^{ij,pq}_{\nu}$ between pairs of antennas numbered $i$ and $j$ respectively, and between
polarizations labeled $p$ and $q$ respectively. We will neglect subtleties associated with polarized
imaging, so the Stokes-$I$ visibilities are constructed from the sum of the $xx$ and $yy$
correlations such that $V^{ij}_{\nu} = (V^{ij,xx}_{\nu}+V^{ij,yy}_{\nu})/2$.  If the antennas are
separated by the baseline $\vec b_{ij}$, and $A_\nu(\hat r)$ describes an antenna's response to the
incident Stokes-$I$ radiation (here assumed to be the same for each antenna), then
\begin{equation}\label{eq:basic-imaging}
    V^{ij}_\nu = \int_\text{sky}
                 A_\nu(\hat r) I_\nu(\hat r)
                 \exp\bigg(2\pi i \hat r\cdot\vec b_{ij}/\lambda\bigg) \,\d\Omega \, .
\end{equation}
Constructing an image from the output of a radio interferometer consists of estimating $I_\nu(\hat
r)$ given the available measurements $V^{ij}_\nu$.

@@

Naively one might attempt to solve Equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations, then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the OVRO-LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit complex
floating point number).  This matrix equation is therefore prohibitively large, and solving
Equation~\ref{eq:basic-imaging} by means of discretization is usually impossible although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

@@

On the other hand, transit telescopes can take advantage of a symmetry in
Equation~\ref{eq:basic-imaging} that greatly reduces the amount of computer time required to image
the full-sky with exact incorporation of widefield imaging effects. This technique, called $m$-mode
analysis, also obviates the need for gridding and mosaicing. Instead the entire sky is imaged in one
coherent synthesis imaging step.  We will briefly summarize $m$-mode analysis below, but the
interested reader should consult \citet{2014ApJ...781...57S, 2015PhRvD..91h3514S} for a complete
derivation.

@@

The power of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$.  By splitting the
equation into $N$ independent blocks, the number of floating point operations required to operate on
the full matrix is reduced by a factor of $N^2$. This computational savings is what makes this
matrix algebra approach to interferometric imaging feasible.

@@

However, usually one will find that $\b B$ is not full-rank and hence $\b B^*\b B$ is not an
invertible matrix. For example, an interferometer located in the northern hemisphere will never see
a region of the southern sky centered on the southern celestial pole. The $m$-modes contained in the
vector $\b v$ must contain no information about the sky around the southern celestial pole, and
therefore the act of multiplying by $\b B$ must destroy some information about the sky. The
consequence of this fact is that $\b B$ must have at least one singular value that is equal to zero.
It is then trivial to show that $\b B^*\b B$ must have at least one eigenvalue that is equal to
zero, which means it is not an invertible matrix.

@@

Tikhonov regularization requires the observer to pick the value of $\varepsilon$. If $\varepsilon$
is too large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. However if $\varepsilon$ is too small then the problem will be
poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\varepsilon$ therefore requires understanding the trade-off between the two norms.

@@

@@

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging (see Figure~\ref{fig:antenna-layout}).  Of these 288 antennas, 251 are contained
within a 200 m diameter core, 32 are placed outside of the core in order to extend the maximum
baseline length to $\sim$1.5~km, and the final 5 antennas are equipped with noise-switched front
ends for calibrated total power measurements of the global sky brightness.  These antennas are used
as part of the LEDA experiment \citep{2017arXiv170909313P} to measure the global signal of 21~cm
absorption from the cosmic dawn. In the current configuration 32 antennas (64 signal paths) from the
core are disconnected from the correlator in order to accommodate the 32 antennas on longer
baselines. A final stage of construction will involve 64 additional antennas installed on long
baselines out to a maximum length of 2.6~km.

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
During this time the OVRO-LWA operated as a zenith-pointing drift scanning interferometer.  The
correlator dump time was selected to be 13 seconds such that the correlator output evenly divides a
sidereal day. Due to the computational considerations presented in \S\ref{sec:mmode-analysis}, eight
24~kHz channels are selected for imaging from this dataset: 36.528~MHz, 41.760~MHz, 46.992~MHz,
52.224~MHz, 57.456~MHz, 62.688~MHz, 67.920~MHz, and 73.152~MHz.

@@

Cyg~A and Cas~A are -- by an order of magnitude -- the brightest point-like radio sources in the
northern hemisphere. Therefore the optimal time to solve for the interferometer's gain calibration
is when these sources are at high elevations.  The antenna complex gains are measured from a 22
minute track of data when Cyg~A and Cas~A are at high elevations. The gains measured in this way are
then used to calibrate the entire 28 hour dataset. The calibration sky model consists only of Cyg~A
and Cas~A. The model flux of Cyg~A is set to the \citet{1977A&A....61...99B} spectrum while the flux
of Cas~A is measured from the data itself (using a preliminary calibration solved for with a
fiducial Cas~A spectrum).

@@

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/beam/beam}
    \caption{
        Empirical fits to the OVRO-LWA Stokes-I primary beam (the response of the $x$- and
        $y$-dipoles has been summed) at three frequencies: 36.528 MHz (left panel), 52.224 MHz
        (middle panel), and 73.152 MHz (right panel). The source tracks used to measure the beam
        model are overlaid. From north to south, these tracks correspond to Cas~A, Cyg~A, Per~B,
        Tau~A, Vir~A, Her~A, 3C~353, and Hya~A.  The fitting process is described in
        \S\ref{sec:beam}, and residuals for Cyg~A and Cas~A are in Figure~\ref{fig:scintillation}.
    }
    \label{fig:beam}
\end{figure*}

@@

We measure the flux of several bright sources (Cyg~A, Cas~A, Tau~A, Vir~A, Her~A, Hya~A,
Per~B\footnote{
    At 74~MHz, Per~B is about two times brighter than Per~A.
}, and
3C~353) as they pass through the sky, and then fit a beam model composed of Zernike polynomials to
those flux measurements. We select the basis functions to have the desired symmetry ($Z_0^0$,
$Z_2^0$, $Z_4^0$, $Z_4^4$, $Z_6^0$, $Z_6^4$, $Z_8^0$, $Z_8^4$, and $Z_8^8$) and the beam amplitude
at zenith is constrained to be unity. See Figure~\ref{fig:beam} for an illustration of a fitted beam
model at several frequencies. This process is repeated for each frequency channel. Residuals for
Cyg~A and Cas~A can be seen in Figure~\ref{fig:scintillation}.

@@

The largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular scale of Cyg~A is
$\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines at 73~MHz, the OVRO-LWA
marginally resolves both sources. A resolved source model is needed for both sources. We fit a
self-consistent resolved source model to each source. This is performed by minimizing the variance
within an aperture located on each source after peeling. By phasing up a large number of
integrations before imaging (at least 1 hour) it is possible to smear out the contribution of the
rest of the sky.  We then use NLopt's Sbplx routine \citep{nlopt, sbplx} to vary the parameters in a
source model until the variance within the aperture is minimized. Cyg~A is modeled with 2 Gaussian
components, while Cas~A is modeled with 3~Gaussian components. Armed with these  source models, both
sources can be peeled from the dataset to the required dynamic range.

@@

Other bright sources -- namely Vir~A, Tau~A, Her~A, Hya~A, Per~B, and 3C~353 -- are also removed
from the visibilities prior to imaging. Because these sources are much fainter than Cyg~A and Cas~A,
we do not need resolved source models to be able to remove these sources from the visibilities
without residual sidelobes contaminating the image.

@@

The Sun can be trivially removed from any map of the sky by constructing the map using only data
collected at night. A map of the entire sky can be obtained by using observations spaced 6 months
apart.  However the dataset used in this paper consists of 28 consecutive hours. Therefore we
attempt to suppress the contribution of the Sun to the maps with the goal of suppressing its
sidelobes.  The Sun is well-resolved by the OVRO-LWA and hence a detailed source model is needed. In
fact the optical depth $\tau=1$ surface of the Sun changes with frequency and as a consequence a new
model is needed at each frequency. While we could fit a limited number of Gaussian components to
Cyg~A and Cas~A, this is insufficient for the Sun.  Additionally while most astronomical sources at
these frequencies have negative spectral indices, the Sun has a positive spectral index. Therefore
more care will need to be taken in subtracting the Sun at higher frequencies than at lower
frequencies.

@@

@@

>>>>

\newcommand{\nrao}{
    \affiliation{
        National Radio Astronomy Observatory,
        P.O. Box O,
        Socorro, NM 87801
    }
}

@@

\newcommand{\swinburne}{
    \affiliation{
        Centre for Astrophysics \& Supercomputing,
        Swinburne University of Technology,
        PO Box 218,
        Hawthorn, VIC 3122, Australia
    }
}

\newcommand{\unm}{
    \affiliation{
        Department of Physics and Astronomy,
        University of New Mexico,
        Albuquerque, NM 87131
    }
}

@@

\newcommand{\nvidia}{
    \affiliation{
        NVIDIA Corporation,
        2701 San Tomas Expressway,
        Santa Clara, CA 95050
    }
}

@@

% alphabetical by last name
\author{Benjamin~R.~Barsdell} \harvard \nvidia
\author{Stephen~A.~Bourke}    \astrocaltech
\author{Steven~W.~Ellingson}  \vt
\author{Jayce~Dowell}         \unm
\author{Hugh~Garsden}         \harvard
\author{Lincoln~J.~Greenhill} \harvard
\author{Jonathon~Kocz}        \astrocaltech
\author{T.~Joseph~W.~Lazio}   \jpl
\author{Daniel~C.~Price}      \harvard \berkeley \swinburne
\author{Frank~K.~Schinzel}    \nrao \unm
\author{Gregory~B.~Taylor}    \unm
\author{David~P.~Woody}       \ovro

@@

\begin{abstract}
    A host of new low-frequency radio telescopes seek to measure the 21-cm transition of neutral
    hydrogen from the early universe.  These telescopes, and the measurements they make, hold
    promise to directly probe star and galaxy formation at redshifts $20 \gtrsim z \gtrsim 7$, but
    are limited by the dynamic range they can achieve against foreground sources of low-frequency
    radio emission. Consequently, there is a growing demand for modern, high-fidelity maps of the
    sky at frequencies below 200 MHz for use in foreground modeling and removal.  We describe a new
    widefield imaging technique for drift-scanning interferometers, Tikhonov-regularized $m$-mode
    analysis imaging.  This technique constructs images of the entire sky in a single synthesis
    imaging step with exact treatment of widefield effects.  We describe how the CLEAN algorithm can
    be adapted to deconvolve maps generated by $m$-mode analysis imaging.  We demonstrate
    Tikhonov-regularized $m$-mode analysis imaging using the Owens Valley Long Wavelength Array
    (OVRO-LWA) by generating 8 new maps of the sky north of $\delta=-30^\circ$ with 15 arcmin
    angular resolution and at frequencies evenly spaced between 36.528~MHz and 73.152~MHz. These
    maps are a 10-fold improvement in angular resolution over existing full-sky maps at comparable
    frequencies, which have angular resolutions $\ge 2^\circ$. The thermal noise in each map is
    $\sim$800 mJy/beam -- comparable to the confusion limit. However, each map is constructed
    exclusively from interferometric observations and does not represent the globally averaged sky
    brightness. Therefore future improvements will incorporate total power radiometry, and
    additional bandwidth and integration time to improve the thermal noise.  The angular resolution
    can also be further improved with the planned expansion of the OVRO-LWA to 2.6~km baselines.
    These maps serve as a first step on the path to the use of more sophisticated foreground filters
    in 21-cm cosmology incorporating the measured angular and frequency structure of all foreground
    contaminants.
\end{abstract}

@@

At redshifts $20 \gtrsim z \gtrsim 7$ the 21-cm hyperfine structure line of neutral hydrogen is
expected to produce a 10 to 100 mK perturbation in the Cosmic Microwave Background (CMB) spectrum
\citep{2006PhR...433..181F, 2012RPPh...75h6901P}. The amplitude of this perturbation on a given
line-of-sight is a function of the neutral fraction of hydrogen, the baryon overdensity, the spin
temperature relative to the CMB temperature at the given redshift, and the line-of-sight peculiar
velocity of the gas.  The spatial power spectrum of this perturbation is thought to be dominated by
inhomogeneous heating of the IGM at $z\sim 20$ \citep{2014MNRAS.437L..36F}, and by growing ionized
bubbles during the EoR at $z\sim 7$ where a detection can constrain the ionizing efficiency of early
galaxies, the UV photon mean-free-path, and the minimum halo mass that can support star formation
\citep{2015MNRAS.449.4246G}.

Current 21-cm cosmology experiments can be broadly separated into two classes: global signal
experiments that are aiming to detect the spectral signature of the cosmologically redshifted 21-cm
transition after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21-cm perturbations.  Ongoing global signal experiments include EDGES
\citep{2010Natur.468..796B, 2017ApJ...847...64M}, LEDA \citep{2017arXiv170909313P}, BIGHORNS
\citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2
\citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

Just as for CMB experiments, foreground removal or suppression is an essential component of both
classes of 21-cm cosmology experiments. The brightness temperature of the galactic synchrotron
emission at high galactic latitudes is measured by \citet{2008AJ....136..641R} as
\begin{equation}
    T \sim 300\,{\rm K} \times \left(\frac{\nu}{150\,{\rm MHz}}\right)^{-2.5}\,.
\end{equation}
Therefore experiments conservatively need to achieve 5 orders of dynamic range against this
foreground emission before the cosmological signal can be measured. Current foreground removal
methods (for example, \citealt{2012ApJ...756..165P}) rely on the assumption that the foreground
emission is spectrally smooth. The low-frequency radio sky is composed of several components:
galactic synchrotron emission, supernova remnants, radio galaxies, free-free emission and absorption
from \ion{H}{2} regions, and a confusing background of radio sources.  Ideally a foreground removal
strategy should be informed by the measured spatial structure and frequency spectrum of all
foreground components. For instance, CMB experiments typically construct several maps at several
frequencies to enable component separation.  At low frequencies, this possibility is limited by the
availability of suitable high-fidelity sky maps on angular scales ranging from tens of degrees to
arcminutes.

@@

The Global Sky Model (GSM) \citep{2008MNRAS.388..247D} is currently the most commonly used
foreground model. The GSM is a non-parametric interpolation of various maps between 10 MHz and 100
GHz. However the majority of information contained in the GSM is derived at frequencies $>1.4$ GHz
where the majority of the modern, high-fidelity input maps are located. At 408~MHz the venerable
Haslam map \citep{1981A&A...100..209H, 1982A&AS...47....1H} covers the entire sky at $1^\circ$
resolution.  Below 408~MHz the GSM uses 3 input sky maps. \citet{2017MNRAS.464.3486Z} construct an
improved GSM with 5 maps below 408~MHz, and \citet{2017MNRAS.469.4537D} use the LWA1 to improve the
GSM with their own sky maps.  However, the GSM generally suffers from low angular resolution ($\sim
5^\circ$) and systematic errors associated with instrumental artifacts in the input maps.  For
instance, \citet{2017MNRAS.469.4537D} report errors of $\pm 50\%$ between the GSM and their own maps
at 74 MHz, which they attribute to the increasing contribution of free-free absorption and
modifications to the synchrotron spectral index at low frequencies.

One challenge in widefield imaging

\todo{briefly touch on m-mode analysis imaging and spherical harmonic imaging??}

This paper presents a series of new low-frequency maps of the sky between 36.528~MHz and 73.152~MHz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21-cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. We anticipate additional ancillary science taking advantage of the combination
of high fidelity and high resolution of these maps, including but not limited to studies of the
cosmic ray emissivity at low frequencies, searches for giant radio galaxies, and constraining the
galactic synchrotron spectrum. The maps will be made freely available online at the Legacy Archive
for Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

@@

The goal of all imaging algorithms is to estimate the brightness of the sky $I_\nu(\hat r)$ in the
direction $\hat r$ and frequency $\nu$.  A radio interferometer measures the visibilities
$V^{ij,pq}_{\nu}$ between pairs of antennas numbered $i$ and $j$ respectively, and between
polarizations labeled $p$ and $q$ respectively. We will neglect subtleties associated with polarized
imaging, so the Stokes-$I$ visibilities are constructed from the sum of the $pp$ and $qq$
correlations such that $V^{ij}_{\nu} = (V^{ij,pp}_{\nu}+V^{ij,qq}_{\nu})/2$.  If the antennas are
separated by the baseline $\vec b_{ij}$, and $A_\nu(\hat r)$ describes an antenna's response to the
incident Stokes-$I$ radiation (here assumed to be the same for each antenna), then
\begin{equation}\label{eq:basic-imaging}
    V^{ij}_\nu = \int_\text{sky}
                 A_\nu(\hat r) I_\nu(\hat r)
                 \exp\bigg(2\pi i \hat r\cdot\vec b_{ij}/\lambda\bigg) \,\d\Omega \, ,
\end{equation}
where the integral runs over solid angle $\Omega$.  Constructing an image from the output of a radio
interferometer consists of estimating $I_\nu(\hat r)$ given the available measurements $V^{ij}_\nu$.

@@

Naively one might attempt to solve Equation~\ref{eq:basic-imaging} by discretizing, and subsequently
solving the resulting matrix equation. If the interferometer is composed of $N_\text{base}$
baselines, and measures $N_\text{freq}$ frequency channels over $N_\text{time}$ integrations, then
the entire data set consists of $N_\text{base}N_\text{freq}N_\text{time}$ complex numbers. If the
sky is discretized into $N_\text{pix}$ pixels then the relevant matrix has dimensions of
$(N_\text{base}N_\text{freq}N_\text{time})\times(N_\text{pix})$. For making single-channel maps with
the OVRO-LWA this becomes a 5 petabyte array (assuming each matrix element is a 64-bit complex
floating point number).  This matrix equation is therefore prohibitively large, and solving
Equation~\ref{eq:basic-imaging} by means of discretization is usually intractable although
\citet{2017MNRAS.465.2901Z} demonstrate this technique with the MITEOR telescope.

@@

Transit telescopes can take advantage of a symmetry in Equation~\ref{eq:basic-imaging} that greatly
reduces the amount of computer time required to image the full-sky with exact incorporation of
widefield imaging effects. This technique, called $m$-mode analysis, also obviates the need for
gridding and mosaicing. Instead the entire sky is imaged in one coherent synthesis imaging step.  We
will briefly summarize $m$-mode analysis below, but the interested reader should consult
\citet{2014ApJ...781...57S, 2015PhRvD..91h3514S} for a complete derivation.

@@

The key advantage of $m$-mode analysis is the block-diagonal structure of
Equation~\ref{eq:m-mode-matrix-equation}. The computational complexity of many common matrix
operations (eg. solving a linear system of equations) is $\mathcal{O}(N^3)$.  By splitting the
equation into $N$ independent blocks, the number of floating point operations required to operate on
the full matrix is reduced by a factor of $N^2$. This computational savings is what makes this
matrix algebra approach to interferometric imaging feasible.

@@

However, usually one will find that $\b B$ is not full-rank and hence $\b B^*\b B$ is not an
invertible matrix. For example, an interferometer located in the northern hemisphere will never see
a region of the southern sky centered on the southern celestial pole. The $m$-modes contained in the
vector $\b v$ must contain no information about the sky around the southern celestial pole, and
therefore the act of multiplying by $\b B$ must destroy some information about the sky. The
consequence of this fact is that $\b B$ must have at least one singular value that is equal to zero.
It then follows that $\b B^*\b B$ must have at least one eigenvalue that is equal to zero, which
means it is not an invertible matrix.

@@

Tikhonov regularization requires the observer to pick the value of $\varepsilon$. If $\varepsilon$
is too large then too much importance is placed on minimizing the norm of the solution and the
least-squares residuals will suffer. Conversely, if $\varepsilon$ is too small then the problem will
be poorly regularized and the resulting sky map may not represent the true sky. Picking the value of
$\varepsilon$ therefore requires understanding the trade-off between the two norms.

@@

\todo{noise covariance matrix?}

@@

The 288 antennas are arranged in a pseudo-random configuration optimized to minimize sidelobes in
snapshot imaging (see Figure~\ref{fig:antenna-layout}).  Of these 288 antennas, 251 are contained
within a 200 m diameter core, 32 are placed outside of the core in order to extend the maximum
baseline length to $\sim$1.5~km, and the final 5 antennas are equipped with noise-switched front
ends for calibrated total power measurements of the global sky brightness.  These antennas are used
as part of the LEDA experiment \citep{2017arXiv170909313P} to measure the global signal of 21-cm
absorption from the cosmic dawn. In the current configuration 32 antennas (64 signal paths) from the
core are disconnected from the correlator in order to accommodate the 32 antennas on longer
baselines. A final stage of construction will involve 64 additional antennas installed on long
baselines out to a maximum length of 2.6~km.

The dataset used in this paper spans 28 consecutive hours beginning at 2017-02-17 12:00:00 UTC time.
During this time the OVRO-LWA operated as a zenith-pointing drift scanning interferometer.  The
correlator dump time was selected to be 13 seconds such that the correlator output evenly divides a
sidereal day. Due to the computational considerations presented in \S\ref{sec:mmode-analysis}, eight
24~kHz channels are selected for imaging from this dataset: 36.528~MHz, 41.760~MHz, 46.992~MHz,
52.224~MHz, 57.456~MHz, 62.688~MHz, 67.920~MHz, and 73.152~MHz.  These particular channels are
chosen due to their location at the exact center of instrumental subbands.

@@

Cyg~A and Cas~A are -- by an order of magnitude -- the brightest point-like radio sources in the
northern hemisphere at resolutions lower than 0.25$^\circ$. Therefore the optimal time to solve for
the interferometer's gain calibration is when these sources are at high elevations.  The antenna
complex gains are measured from a 22 minute track of data when Cyg~A and Cas~A are at high
elevations. The gains measured in this way are then used to calibrate the entire 28 hour dataset.
The calibration sky model consists only of Cyg~A and Cas~A. The model flux of Cyg~A is set to the
\citet{1977A&A....61...99B} spectrum while the flux of Cas~A is measured from the data itself (using
a preliminary calibration solved for with a fiducial Cas~A spectrum).

@@

\todo{mention temperature variations at the front end}

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/beam/beam}
    \caption{
        Empirical fits to the OVRO-LWA Stokes-I primary beam (the response of the $x$- and
        $y$-dipoles has been summed) at three frequencies: 36.528 MHz (left panel), 52.224 MHz
        (middle panel), and 73.152 MHz (right panel). The source tracks used to measure the beam
        model are overlaid. From north to south, these tracks correspond to Cas~A, Cyg~A, 3C~123,
        Tau~A, Vir~A, Her~A, 3C~353, and Hya~A.  The fitting process is described in
        \S\ref{sec:beam}, and residuals for Cyg~A and Cas~A are in Figure~\ref{fig:scintillation}.
    }
    \label{fig:beam}
\end{figure*}

@@

We measure the flux of several bright sources (Cyg~A, Cas~A, Tau~A, Vir~A, Her~A, Hya~A, 3C~123, and
3C~353) as they pass through the sky, and then fit a beam model composed of Zernike polynomials to
those flux measurements. We select the basis functions to have the desired symmetry ($Z_0^0$,
$Z_2^0$, $Z_4^0$, $Z_4^4$, $Z_6^0$, $Z_6^4$, $Z_8^0$, $Z_8^4$, and $Z_8^8$) and the beam amplitude
at zenith is constrained to be unity. See Figure~\ref{fig:beam} for an illustration of a fitted beam
model at several frequencies. This process is repeated for each frequency channel. Residuals for
Cyg~A and Cas~A can be seen in Figure~\ref{fig:scintillation}.

@@

The largest angular scale of Cas~A is $\sim$8~arcmin, and the largest angular scale of Cyg~A is
$\sim$2~arcmin. With a $\sim10$~arcmin resolution on its longest baselines at 73~MHz, the OVRO-LWA
marginally resolves both sources. A resolved source model is needed for both sources. We fit a
self-consistent resolved source model to each source. This is performed by minimizing the variance
within an aperture located on each source after peeling. By phasing up a large number of
integrations before imaging (at least 1 hour) it is possible to smear out the contribution of the
rest of the sky.  We then use a non-linear optimization routine (NLopt Sbplx) \citep{nlopt, sbplx}
to vary the parameters in a source model until the variance within the aperture is minimized. Cyg~A
is modeled with 2 Gaussian components, while Cas~A is modeled with 3~Gaussian components. Armed with
these  source models, both sources can be peeled from the dataset to the required dynamic range.

@@

Other bright sources -- namely Vir~A, Tau~A, Her~A, Hya~A, 3C~123, and 3C~353 -- are also removed
from the visibilities prior to imaging. Because these sources are much fainter than Cyg~A and Cas~A,
we do not need resolved source models to be able to remove these sources from the visibilities
without residual sidelobes contaminating the image.

@@

The Sun can be trivially removed from any map of the sky by constructing the map using only data
collected at night. A map of the entire sky can be obtained by using observations spaced 6 months
apart.  However the dataset used in this paper consists of 28 consecutive hours. Fortunately the Sun
was not active during this period, which could have greatly increased the difficulty involved in
subtracting the Sun.

We attempt to subtract the Sun from the dataset with the goal of suppressing its sidelobes.  The Sun
is well-resolved by the OVRO-LWA and hence a detailed source model is needed. In fact the optical
depth $\tau=1$ surface of the Sun changes with frequency and as a consequence a new model is needed
at each frequency. While we could fit a limited number of Gaussian components to Cyg~A and Cas~A,
this is insufficient for the Sun.  Additionally while most astronomical sources at these frequencies
have negative spectral indices, the Sun has a positive spectral index. Therefore more care will need
to be taken in subtracting the Sun at higher frequencies than at lower frequencies.

@@

\todo{there is only one subsection here, can it be removed?}

@@

\todo{compare with obenberger 2015}
